Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-03-29 ms.tex
more work on low-order corrections section

In the course of this study we use not only \PLANCK~857 GHz, but also all of
the remaining \PLANCK~HFI bands as well as \verb|i100|. To derive the 
absolute zero level offsets that must be applied to each of the five
lowest-frequency \PLANCK~bands, we perform a regression versus the 
\PLANCK~857 GHz map corrected for the absolute zero level offset derived in
 $\S$\ref{sec:zp_abs}. We assume no offset need be applied
to \verb|i100|, which has a zero level tied to H\,\textsc{i} by SFD.
% maybe need to restate the idea that we're deriving zero points by assuming
% that the band of interest must approach zero linearly as the reference
% (in this case 857 ghz) approaches zero.

Inspecting the HFI maps at 100-545 GHz, the need for low-order corrections beyond
a scalar offset were evident. In particular, we noticed the presence of a residual 
Solar dipole. To best recover the amplitudes of the low-order corrections, it is
necessary to have the highest-possible signal in the independent variable of our
regression. For this reason we have used 857 GHz as a reference for the other HFI 
bands, as opposed to the FDS99 predictions or H\,\textsc{i}.

Importantly, we have performed one regression per HFI band to simultaneously fit for the 
zero level offset, the slope relative to 857 GHz, and the residual dipole amplitude.
For each 100-545 GHz HFI band, we restrict tor regions of low column density
(H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ km\,s$^{-1}$)
and fit the following model:

%At 353 GHz, 545 GHz, 857 GHz, we have 
%adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
% by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column density, and
%then by correlating the other \PLANCK~channels against the corrected 857 GHz 
%map. We have not altered the \verb|i100| zero-point in any way. We have also 
%removed the Solar dipole residual according to \cite{planckdust}, by 
%subtracting a dipole pattern oriented toward 
%$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
%Table \ref{table:offs}.

@@

\begin{equation}
\mathcal{I}_{\nu_i, p} = a\mathcal{I}_{857, p} + b\mathcal{D}_{p} + c
\end{equation}

With $p$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low H\,\textsc{i} cut, we have an overconstrained linear model with
three parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improved correlation striking in this scatterplot, 
but the residual dipole also appears visually well-corrected in the 217 GHz 
map itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.

>>>>

In the course of this study we use not only \PLANCK~857 GHz, but also all of
the remaining \PLANCK~HFI bands, as well as \verb|i100|. To derive the 
absolute zero level offsets that must be applied to each of the five
lowest-frequency \PLANCK~bands, we perform a regression versus the 
\PLANCK~857 GHz map corrected for the absolute zero level offset derived in
 $\S$\ref{sec:zp_abs}. We assume no offset need be applied
to \verb|i100|, which has a zero level tied to H\,\textsc{i} by SFD.
% maybe need to restate the idea that we're deriving zero points by assuming
% that the band of interest must approach zero linearly as the reference
% (in this case 857 ghz) approaches zero.

The need for additional low-order corrections beyond a simple scalar offset 
became evident upon inspecting the HFI maps at 100-545 GHz. In particular, we 
noticed the presence of a residual dipole pattern, with an orientation 
consistent with that of the Solar dipole. To most precisely recover the 
amplitude of the residual dipole, it is necessary to have the 
highest-possible S/N in the independent variable of our regression. For 
this reason we have used 857 GHz as a reference for the 100-545 GHz bands, as 
opposed to the FDS99 predictions or H\,\textsc{i}.

We have performed one regression per HFI band to simultaneously 
fit for the zero level offset, the slope relative to 857 GHz, and the residual 
dipole amplitude. For each 100-545 GHz HFI band, we restrict tor regions of low
column density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for 
$-72$$<$$v_{LSR}$$<$$+25$ km\,s$^{-1}$) and fit the following model:

% At 353 GHz, 545 GHz, 857 GHz, we have 
% adopted the zero-point offsets of \cite{planckdust} Table 1, which were 
% derived by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column 
% density, and
%then by correlating the other \PLANCK~channels against the corrected 857 GHz 
%map. We have not altered the \verb|i100| zero-point in any way. We have also 
%removed the Solar dipole residual according to \cite{planckdust}, by 
%subtracting a dipole pattern oriented toward 
%$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
%Table \ref{table:offs}.

@@

\begin{equation} \label{equ:dip}
\mathcal{I}_{\nu_i, p} = a\mathcal{I}_{857, p} + b\mathcal{D}_{p} + c
\end{equation}

With $p$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{\nu_i}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and the constant offset of $\S$\ref{sec:zp_abs} removed, smoothed to 
$1^{\circ}$ resolution. $\mathcal{I}_{\nu_i}$ is the corresponding $1^{\circ}$ 
resolution \PLANCK~HFI map with zodiacal emission and compact sources removed. 
In the context of Equation \ref{equ:dip}, $\nu_i \in$  
\{100, 143, 217, 353, 545\} GHz. $\mathcal{D}$ is a scaling of the Solar dipole
pattern oriented toward $(l, b) = (263.99^{\circ}, 48.26^{\circ})$, with unit 
amplitude. Because $\sim$18,000 pixels satisfy the low H\,\textsc{i} cut, we 
have an overconstrained linear model with three parameters: a, b, and c. $a$ 
represents the slope of \PLANCK~band $\nu_i$ GHz versus \PLANCK~857 GHz 
assuming they are linearly correlated. $b$ represents the scaling of the Solar 
dipole that makes the 857 GHz vs. \PLANCK~band $\nu_i$ correlation most tightly
linear, and $c$ represents the constant offset necessary to make the zero level
 of the band $\nu_i$ map consistent with that of the 857 GHz map. Performing a 
linear least squares fit with uniform weights and iterative outlier rejection, 
we find $a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improved correlation striking in this scatterplot, 
but the residual dipole also appears visually well-corrected in the 217 GHz 
map itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.
